Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB

### 2018-06-25 paper/paper.tex
updated text

@@

	Information processing by neurons relies on the generation and transmission of electrical signals that arise from the biophysics of ion channels, synapses and other voltage-coupled components. Electrophysiological characterization of these mechanisms has allowed for the construction of conductance-based models that can reproduce several features of neuronal and circuit behavior. However, working with conductance-based models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features. Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB". Our approach unites the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB". In our software, models are hierarchical, with components that are named and searchable, permitting high-level programmatic control over all parameters. The simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately. The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics. Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily into the workflow of many researchers.

@@

Nervous systems process and transmit information using electrically excitable membranes. Conductance-based models are a powerful biophysical simplification of an excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}). Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}). These models provide an approachable framework for understanding many salient principles of neuroscience, since they explicitly model cell membranes and ion channels as electrical components in a circuit. However, several challenges remain in working with biophysically-detailed conductance-based neuron models. First, these models can be high-dimensional with many nonlinear differential equations, with several parameters. Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.  In multi-compartment models, membrane potentials in every compartment are coupled to each other. Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++", "FORTRAN" or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".

@@

To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files. This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, without having to re-specify these objects in "MATLAB". In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models. Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax. Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects. "xolotl" comes packaged with hundreds of conductances and synapses, built-in visualization functions, and a graphical user interface (GUI) for real-time manipulation of model parameters. Built-in methods allow the user to easily generate plots of voltage and intracellular calcium time traces, and activation functions of constituent conductances. The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model. The ease-of-use of these tools lends them to pedagogical applications and the rapid prototyping of models. This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and complement other tools being developed in the computational neuroscience community.

@@

"xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research. Our focus was on designing a simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software. Specifically, the software was designed to simulate models of the form

\[ C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \]

@@

\[ I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i} \]

@@

Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \). The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used.

This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research. Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable so that the software can be used as a component of other packages and tools. All parameters of models and activation functions of any channel can be be accessed before running the model. Finally, we designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file. Models created using this simulator appear in the "MATLAB" workspace as native objects, and are fully compatible with the large library of toolboxes that "MATLAB" provides.

@@

\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) that can be used as building blocks to construct model neurons and networks. Compartments represent sections of membrane with a single membrane potential, and can represent either entire cells or parts of cells. Conductances represent populations of ion channels in a compartment that produce transmembrane currents. Synapses connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron. Controllers represent homeostatic mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.

\paragraph{Object-oriented.}
"xolotl" is designed to mirror the nested and hierarchical structure of networks and neurons. Biological neuronal networks are made up of neurons that are connected to each other with synapses. Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior. These conductances can be controlled by a feedback system that can homeostatically regulate their properties. Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons. Each compartment can contain a arbitrary set of conductances, that may be regulated by a set of controllers. All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods. This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing and integrating these models easier.

\paragraph{Automatic type system.}
Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB". This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that do not exist in "MATLAB". A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations. However, this approach is cumbersome, and prone to errors, since the same code needs to be written twice, in two languages. Instead, our automatic type system creates objects in "MATLAB" on-the-fly  from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation.

\paragraph{Automatic hashing and compiling.}
Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable. "xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier. Compiled binaries are named with this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed. This powerful feature enables the user experience to remain entirely within the "MATLAB" workspace, with compilation and selection of the correct binary occurring silently in the background.

@@

Our focus on ease-of-use and fast simulations means some features were intentionally neglected in the design process.

@@

\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}. However, it may be desirable to use other methods under certain conditions. It is possible to introduce new network components that implement other integration schemes, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.

\paragraph{Inefficient tools for handling large networks.} Since every compartment and component in a "xolotl" model is $named$, rather than numbered, it is more suited towards simulating small but complex networks, rather than large but simple networks. While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.

@@

The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current. Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}). The Hodgkin-Huxley model generates action potentials, and, as one of the simplest models of excitable neural mebranes, often serves as a the first model introduced in pedagogical literature (cite some textbooks).

@@

In this section, we demonstrate how to simulate a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism. The schematic in Figure \ref{fig:figurehh} shows how the hierarchical structure of the ion channel populations within the soma is represented in the simulator, and how to create a compartment (here, labelled "HH") and add three populations of ion channels to it. These models of ion channels were obtained from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator. Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage. In the absence of injected current, the model is quiescent. When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}A-B).  The \texttt{plot} function generates a voltage trace, which is colored by the dominant current (Figure \ref{fig:figurehh}A). If the membrane potential is increasing, the strongest instantaneous inward current colors the trace. Conversely, if the membrane potential is decreasing, the strongest outward current colors the trace instead. These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current. This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.

The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (cite). Figure \ref{fig:figurehh}C shows the f-I curve of this model, obtained by repeated integration of the model.  Finally, the built-in \texttt{show} function can plot activation (m) and inactivation (h) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G). These plots revleal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential. In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides a rich visualization of the dynamics of the model.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
	\caption{Simulating a voltage-clamp experiment. (A) Cartoon of a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998} with experimentally-fixed voltage. (B) Structure of \texttt{xolotl} object in A. (C) Code snippet depicting integration under voltage clamp. (D-E) Current response to steps in voltage from a holding potential of $V_m = -60$ mV. (F) Current-voltage relation of the steady-state current ($t = 400$ ms) indicating a reversal potential of $E = -80$ mV and no inactivation. (G) Conductance-voltage relation at steady-state takes the form of a sigmoid. (H) Sigmoids $m$ fit to the model as $m^n$ data indicating that $n=4$ is the best fit. (I) Goodness of fit vs. exponent $n$, suggesting $n=4$ as the best fit to the data.}
	\label{fig:figureclamp}
\end{figure}

"xolotl" can recapitulate the results of voltage clamp experiments (\cite{turrigianoSelectiveRegulationCurrent1995, swensenMultiplePeptidesConverge2000, swensenModulatorsConvergentCellular2001, destexheDynamicClampPrinciplesApplications2009}). Figure \ref{fig:figureclamp} displays steps in the procedure to clamp the membrane potential of a cell with delayed rectifier potassium conductance. During an \textit{in-vitro} experiment, confounding currents would be pharmacologically-blocked and two-electrode voltage clamp used to record tail currents at fixed membrane potential (\cite{connorInwardDelayedOutward1971, connorVoltageClampStudies1971}).

A single-compartment model with a delayed-rectifier conductance (\cite{liuModelNeuronActivitydependent1998}) is simulated at stepped membrane potentials. The "integrate" function outputs the membrane potential and current as time series. Currents under voltage clamp approach the steady-state holding current (Figure \ref{fig:figureclamp}D-E). The current-voltage relation is the steady-state current over the clamped voltage, and the effective conductance is the derivative of that relation (Figure \ref{fig:figureclamp}F-G). Since the effective conductance is the product of the maximal conductance and the gating variables (\cite{dayanTheoreticalNeuroscience2001, turrigianoSelectiveRegulationCurrent1995}) and the tail current is monotonically-increasing with time under voltage clamp, the current can be represented as non-inactivating. Fitting a sigmoid to various powers yields a model for the current dynamics (Figure \ref{fig:figureclamp}H-I). These figures describe graphically the theoretical underpinnings of current analysis through voltage clamp and can serve as an effective pedagogical tool for computational and quantitative neuroscience.

>>>>

\newcommand{\dd}[1]{\mathrm{d}#1}

@@

	Conductance-based models of neuron dynamics based on Hodgkin-Huxley formalism have been used extensively in computational neuroscience. However, working with these models continues to be a challenge due to their high dimensionality and large number of parameters, hindering intuition of their dynamical features. Here, we present a neuron and network simulator built on a novel automatic type system that binds class templates written in "C++" to object-oriented code in "MATLAB". Our approach builds on the tradition of uniting the speed and power of "C++" code with the ease-of-use and rich libraries of scientific programming languages like "MATLAB". In our software, models are hierarchical, with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters. Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately. The simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be easily extended to include arbitrary mechanisms and dynamics. Finally, the simulator is written in a modular fashion and has been released under a permissive free software license, enabling it to be integrated easily into the workflow of many researchers.

@@

Nervous systems process and transmit information using electrically excitable membranes. Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}). Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}). These models provide an approachable framework for understanding many salient principles of neuroscience, since they explicitly model cell membranes and ion channels as electrical components in a circuit. However, several challenges remain in working with biophysically-detailed conductance-based neuron models. First, these models can be high-dimensional with many nonlinear differential equations, with several parameters. Second, many or all equations in these models can be strongly coupled through variables like the membrane potential.  In multi-compartment models, membrane potentials in every compartment are coupled to each other. Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like "C++", "FORTRAN" or "hoc" can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like "Python", "Julia" or "MATLAB".

@@

To overcome these design limitations, we have developed a novel automatic type system, "cpplab", which binds "MATLAB" code to "C++" header files. This architecture automatically creates objects in "MATLAB" that represent the underlying object-oriented "C++" code, without having to re-specify these objects in "MATLAB". In this paper, we introduce "xolotl", an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models. Models can be easily constructed from network building blocks in a few lines of "MATLAB" code using a hierarchical and intuitive syntax. Since models in the "MATLAB" workspace are automatically linked to models in the "C++" implementation, configuring these objects in "MATLAB" transparently configures the underlying "C++" objects. "xolotl" comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a graphical user interface (GUI) for real-time manipulation of model parameters. The GUI permits real-time manipulation of any network parameter and an immediate visualization of the effect of that parameter on arbitrary dynamical features of the model. The ease-of-use of these tools lends them to pedagogical applications and the rapid prototyping of models. This tool aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.

@@

"xolotl" is designed to be easy-to-use and richly featured while being fast enough to use in everyday research. Our focus was on designing an approchable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software. Specifically, the software was designed to simulate models of the form

\begin{equation}
C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \label{eq:1}
\end{equation}

@@

\begin{equation}
I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i}  \label{eq:2}
\end{equation}

@@

Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \). The software uses integration schemes that have been specifically developed to solve equations of this form (cite D\&A, Hines, others), though other schemes can be used if desired.

This software is designed to be used from within "MATLAB", a scientific programming language common amongst neuroscientists and engineers for pedagogy and research. Our goal was to make "xolotl" completely usable entirely from within MATLAB, and fully scriptable. Models created using this simulator appear in the "MATLAB" workspace as native objects, and are thus fully compatible with the large library of toolboxes that "MATLAB" provides, allowing so the software to be used as a component of other packages and tools. All parameters of models and activation functions of any channel can be be inspected before running the model. We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by changing a few lines in a template file. Finally, "xolotl" is designed to be fully auditable, and comes with several tools to verify model and parameter integrity and aid reproducibility.

@@

\paragraph{A rich library of network components.} "xolotl" comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and controllers) that can be used as building blocks to construct model neurons and networks. Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components; and can represent either entire cells or parts of cells. Conductances represent populations of ion channels in a compartment that produce transmembrane currents. Synapses connect two compartments together by introducing a current in the post-synaptic neuron that can depend on the voltage of the presynaptic neuron. Controllers represent intracellular mechanisms that integrate a Calcium-dependent signal from the compartment they are in and  change the properties of a conductance or synapse in response.

\paragraph{Object-oriented.}
"xolotl" is designed to mirror the nested and hierarchical structure of networks and neurons. Biological neuronal networks are made up of neurons that are connected to each other with synapses. Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior. These conductances can be controlled by intracellular regualtory system that can change their properties. Similarly, a "xolotl" model can contain a set of compartments that can represent individual neurons. Each compartment can contain a arbitrary set of conductances, that may be regulated by a set of controllers. All objects in the "xolotl" model tree (compartments, conductances, etc.) are bonafide "MATLAB" objects that have their own type, properties and associated methods. This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing and integrating these models easier.

\paragraph{Automatic type system.}
Our approach to circumvent the tradeoff between high-performance but hard to use languages like "C++" and richly featured but slow languages like "MATLAB" was to construct an automatic type system that links object oriented code in "C++" to object oriented code in "MATLAB". This architecture lets us construct the core of the simulator in "C++", leveraging features of "C++" like pointers that are not commonly used in "MATLAB". A rudimentary way to make this "C++" code useable in "MATLAB" would be to re-write that code in "MATLAB" so that "MATLAB" objects can be bound their "C++" implementations. However, this approach is cumbersome, and prone to errors, since the same code needs to be written twice, in two languages. Instead, our automatic type system creates objects in "MATLAB" on-the-fly  from "C++" class specifications, obviating the need to rewrite code in "MATLAB" while preserving a tight coupling between objects in the "MATLAB" workspace and their "C++" implementation. The key advantages of this approach are that it makes developing new code much easier, and simplifies the task of constructing complex frameworks that span these two languages.

\paragraph{Automatic hashing and compiling.}
Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable. "xolotl" solves this problem by hashing the "C++" header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its "hash"). Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed. This powerful feature enables the user experience to remain entirely within the "MATLAB" workspace, with compilation and selection of the correct binary occurring silently in the background.

@@

Our focus on ease-of-use and speed means some features were intentionally neglected in the design process.

@@

\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}. However, it may be desirable to use other methods under certain conditions. It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new "C++"  code, and is still restricted to integration schemes that assume a fixed step size.

\paragraph{Inefficient tools for handling large networks.} "xolotl" was designed to work with small but complex networks and models, where every compartment and component is $named$, rather than numbered. It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks. While the software $can$ integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task.

@@

The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current. Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane, for example, by injecting current, could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}). As one of the simplest models of excitable neural mebranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (cite some textbooks).

@@

In this section, we demonstrate how to simulate a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model, and gain insight into the underlying biophysical mechanism. The schematic in Figure \ref{fig:figurehh} shows how the hierarchical structure of the ion channel populations within the soma is represented in the simulator, and how to create a compartment (here, labelled "HH") and add three populations of ion channels to it. These models of ion channels were obtained from \cite{liuModelNeuronActivitydependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator. Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage. In the absence of injected current, the model is quiescent. When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}A-B).  The \texttt{plot} function generates a voltage trace, which is colored by the dominant current (Figure \ref{fig:figurehh}A). Colors in the voltage trace indicate the strongest instantaneous inward current when the voltage is increasing, and strongest instantaneous outward current when the voltage is decreasing. These built-in features reveal that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current, but that the leak current contributes to depolarization following an action potential. This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.

Integrating the model returns the voltage time series for every compartment:

\begin{equation}
{\tt V = x.integrate;}
\end{equation}

The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (cite). Figure \ref{fig:figurehh}C shows the f-I curve of this model, obtained by repeated integration of the model.  Finally, the built-in \texttt{show} function can plot activation ($m$) and inactivation ($h$) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G). These plots reveal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential. In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides a rich visualization of the dynamics of the model.

@@

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
	\caption{Simulating a voltage-clamp experiment. The cartoon shows a cell with delayed rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) that is being recorded from in two-electrode voltage clamp. The code snippet sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it. (A) Voltage steps that the cell is clamped to. (B) Current injected into the cell to clamp it to the voltages in (A). (C) Asymptotic clamped current vs. clamped voltage for this cell. (D) Accounting for the reversal potential of Potassium ions yeilds the conductance-voltage curve of this channel type. (E) Normalized conductance-voltage curves, with sigmoid fits with various exponents. (F) An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type. }
	\label{fig:figureclamp}
\end{figure}

Voltage clamping is a experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (cite D\&A). Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp. By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized. (insert all those citations as needed).

"xolotl" can reproduce such a voltage clamp experiment $in silico$. Figure \ref{fig:figureclamp} illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.  Integrating the model yields the current required to clamp the cell at that voltage. Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivitydependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel. First, the cell is clamped to a number of different voltages (Fig. \ref{fig:figureclamp}A) and the resultant clamping currents are measured by integrating the model (Fig. \ref{fig:figureclamp}B). The asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig. \ref{fig:figureclamp}C), since the open probabilities of the channel are functions of the membrane voltage. Assuming the reversal potential is known, Eq. \eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig. \ref{fig:figureclamp}D). Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activiation function of the ion channel population (Fig. \ref{fig:figureclamp}E-F). "xolotl" can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.
